---
title: "ggplot2"
author: "Michael Gebhardt"
date: "September 15, 2015"
output: html_document
---

# ggplot2

```{r}
library("ggplot2")
```

```{r plot1}
p <- ggplot(data = research, mapping = aes(x = pdfDownloadsCount,
                                           y = wosCountThru2011)) +
  geom_point()
p
```

In ggplot2, you are mapping your data onto layers on the plot
The '+' sign is relatively equivalent to the pipe command from dplyr (%>%)
```{r plot1}
p <- ggplot(data = research, aes(x = pdfDownloadsCount,
                  # Can remove mapping, since it's default
                                 y = wosCountThru2011)) +
  geom_point(aes(color = journal))
p
```

```{r}
p <- ggplot(data = research, aes(x = pdfDownloadsCount,
                                 y = wosCountThru2011)) +
  geom_point(aes(size = authorsCount))
p
```

```{r}
p <- ggplot(data = research, aes(x = pdfDownloadsCount,
                                 y = wosCountThru2011,
                                 color = journal)) +
  geom_point() +
  geom_smooth()
p
```
Can omit aes and then it will ignore differences in data and color everything one color. 
'aes' will apply aesthetics based on the data itself (data dependent)

# Challenge Lesson 15

Create a scatter plot with daysSincePublished mapped to the x-axis and wosCountThru2011 mapped
to the y-axis. Include a loess fit of the data. Set the transparency level (alpha) of the points
to 0.5 and color the points according to the journal where the article was published. Make the 
loess curve red.

```{r}
p <- ggplot(research, aes(x = daysSincePublished,
                          y = wosCountThru2011)) +
  geom_point(aes(color = journal), alpha = 0.5) +
  geom_smooth(color = "red")
p
```
The plot has a single curve plotted in red, showing a slight positive correlation between citation number and age
Also the points are semi transparent, so stacked points are darker

### USING SCALES

```{r}
p <- ggplot(data = research, aes(x = pdfDownloadsCount,
                                 y = wosCountThru2011)) +
  geom_point(aes(color = journal)) +
  geom_smooth()
p + scale_x_log10() + scale_y_log10()
```
Since we're doing a log scale, any values with a 0 will throw an error message; can intercept:

```{r}
p <- ggplot(data = research, aes(x = log10(pdfDownloadsCount + 1),
                                 y = log10(wosCountThru2011 + 1))) +
  geom_point(aes(color = journal)) +
  geom_smooth() +
  scale_x_continuous(breaks = c(1, 3), labels = (c(10,1000))) +  
  # This keeps the x axis labels in check; since without this value the axis labels disappear
  # Also keeps the axis scaled appropriate for log10 scale
  scale_y_continuous(breaks = c(1, 3), labels = c(10, 1000),
                     limits = c(1, 3))
p
```

## Different Color Options

```{r}
p + scale_color_grey()
```

```{r}
p + scale_color_manual(values = c("red", "green", "blue", "orange", "pink", "yellow", "purple"))
```

```{r}
library("RColorBrewer")
display.brewer.all(type = "qual")
```

```{r}
p + scale_color_brewer(palette ="Set1", 
                       labels = 1:7, name = "PLOS")
```

## GGPLOT CHALLENGE

Update the plot to use a square root transfomration instead of log10.  Also, color the points using the ColorBrewer palette "Accent".

```{r}
p <- ggplot(data = research, aes(x = sqrt(pdfDownloadsCount),
                                 y = sqrt(wosCountThru2011))) +
  geom_point(aes(color = journal)) +
  geom_smooth()
p
p + scale_color_brewer(palette = "Accent")
```

Can also remove the 'sqrt' transformations in the aes command and do :
  'p + scale_x_sqrt' and '+ scale_y_sqrt'
  
### Using facets to make subplots

```{r}
p <- ggplot(data = research, aes(x = sqrt(pdfDownloadsCount),
                                 y = sqrt(wosCountThru2011))) +
  geom_point(aes(color = journal)) +
  geom_smooth()
p + facet_wrap(~journal)
```

Items in aes() are forced to all layers on the plot?

Using 'facet_grid'

```{r}

research <- mutate(research, immuno = grepl("Immunology", plosSubjectTags))
p + facet_grid(journal~immuno)
```
This gives and error message because we've changed the research data (added immuno), but original call to ggplot used different data frame.
```{r}
research <- mutate(research, immuno = grepl("Immunology", plosSubjectTags))
p <- ggplot(data = research, aes(x = sqrt(pdfDownloadsCount),
                                 y = sqrt(wosCountThru2011))) +
  geom_point(aes(color = journal)) +
  geom_smooth()
p + facet_grid(journal~immuno)
```

### Using different geoms

```{r}
p <- ggplot(data = research, aes(x = journal,
                                 y = sqrt(wosCountThru2011))) +
  geom_boxplot()
p
```

### Making a barplot

```{r}
# Get data
tweets_per_journal <- research %>% group_by(journal) %>%
  summarize(number_articles = n(), 
            tweets_mean = mean(backtweetsCount),
            tweets_SEM = sd(backtweetsCount) / sqrt(number_articles))
tweets_per_journal
```

```{r}
tweets_bar <- ggplot(tweets_per_journal, aes(x = journal,
                                             y = tweets_mean)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = tweets_mean - tweets_SEM, ymax = tweets_mean + tweets_SEM), width = 0.1)+
  geom_text(aes(label = number_articles), hjust = 0, vjust = -0.1)
tweets_bar
```
'hjust' = horizontal justification labels
'vjust' = vertical justification for labels

# Challenges for Lesson 18
Challenge 1:

Modify dplyr code above to calculate the mean, SEM and sample size of the number of
article tweets per journal and per year. Use 'facet_wrap' to make a separate subplot
per year. 

```{r}
# Modify data frame:
tweets_challenge <- research %>% group_by(journal, year) %>%
  summarize(num = n(), 
            mean = mean(backtweetsCount),
            SEM = sd(backtweetsCount) / sqrt(num))
tweets_challenge
# Now the plots:
tweets_bar_chall <- ggplot(tweets_challenge, aes(x = journal,
                                             y = mean)) +
  geom_point(stat = "identity") +
  geom_errorbar(aes(ymin = mean - SEM, ymax = mean + SEM), width = 0.1)
tweets_bar_chall
tweets_bar_chall + facet_wrap(~year)
```

Challenge 2: Visualizing a single distribution
The geoms 'geom_histogram' and 'geom_density' can be used to create histograms
and density plots, respectively. Using these geoms, visualize the distribution
of 2011 citations (wosCountThru2011). Compare the raw distribution to log10 
and square root transformations.

```{r}
??
```


## Customizing the Plots

```{r}
tweets_bar_chall + labs(title = "Mean tweets per journal per year", x = "Journal", y = "Mean Number of Tweets") + facet_wrap(~year) + 
  theme_classic()
```

```{r}
tweets_bar_chall + labs(title = "Mean tweets per journal per year", x = "Journal", y = "Mean Number of Tweets") + facet_wrap(~year) + 
  theme_bw()
```

Can set default themes by using 
'theme_set(theme_bw())'